#CODEBOOK

This is the codebook for the data analysis performed. 

## Step 01: Merges the training and the test sets to create one data set.
All the anaylsis is performed in *run_analysis.R*. For merging the datsets, the following procedure is followed:

- Load all the data files using read.table()

- The data sets are merged in the "wide" sense- The training data set sits on top of the test data set. 

- After merging the datasets, columns indicating the subject ID, the activity ID, and whether a particular observation (row in the dataframe) belongs to the training set or the test set are also added.

- This results in a merged dataset called "rawdata" with dimensions of: 10299 obs. of 564 variables. 

- Please see the equivalent R code:
```{r}
library(dplyr)

## Load the subject IDs
sID_train <- read.table ("subject_train.txt")
sID_test <- read.table ("subject_test.txt")
subjID <- rbind(sID_train, sID_test)
subjID <- rename (subjID, subjID = V1)

## load the data sets
data_train <- read.table("X_train.txt")
data_test <- read.table("X_test.txt")

## load the activity labels for the data sets
actlabel_train <- read.table("y_train.txt")
actlabel_test <- read.table("y_test.txt")
activityID <- rbind(actlabel_train, actlabel_test)
activityID <- rename (activityID, activityID = V1)

## Merge the data values
rawdata <- rbind(data_train, data_test)

## Create a column vector for indicating which observations belong to training data set and which to test dataset
dataset <- rep(c("Training","Test"), c(dim(data_train)[1], dim(data_test)[1]))

## Merge with activity ID
rawdata <- cbind(activityID, rawdata)

## Merge with Subject ID 
rawdata <- cbind(subjID, rawdata)

## Merge with Data Set Type 
rawdata <- cbind(dataset, rawdata)

## Load feature information
feature <- read.table("features.txt")
```

## Step 02: Extracts only the measurements on the mean and standard deviation for each measurement. 
For this step, those columns are selected which hold the values of the mean and standard deviation of each measurment. It stores these columns in a new data frame called *data*.
```{r}
data <- select(rawdata, dataset, subjID, activityID, V1, V2, V3, V4, V5, V6, V41
               , V42, V43, V44, V45, V46, V81, V82, V83, V84, V85, V86, V121, 
               V122, V123, V124, V125, V126, V161, V162, V163, V164, V165, V166,
               V201, V202, V214, V215, V227, V228, V240, V241, V253, V254, V266,
               V267, V268, V269, V270, V271, V345, V346, V347, V348, V349, V350, 
               V424, V425, V426, V427, V428, V429, V503, V504, V516, V517, V529, 
               V530, V542, V543 )
```

##Step 03: Uses descriptive activity names to name the activities in the data set
Before renaming the activities from their numeric mnemonic to real-life name for example: Walking, the class of the columns is coerced to numeric. 
```{r}
data$activityID <- as.numeric(as.character(data$activityID))
data$subjID <- as.numeric(as.character(data$subjID))
data$dataset <- as.character(data$dataset)

data$activityID[data$activityID == 1] <- "Walking"
data$activityID[data$activityID == 2] <- "WalkingUpStairs"
data$activityID[data$activityID == 3] <- "WalkingDownStairs"
data$activityID[data$activityID == 4] <- "Sitting"
data$activityID[data$activityID == 5] <- "Standing"
data$activityID[data$activityID == 6] <- "Laying"
```
##Step 04: Appropriately labels the data set with descriptive variable names. 
The variable names given to the measurements are the same as given in the initial data files, except for the addition of the mean and sd suffix. That is: 

The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag). 

Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals). 

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.

**Stating again: symbols in this analysis differ from the original description of the measurements by the addition of a mean and sd suffix **.
```{r}
data <- rename (data, tBodyAccX_mean = V1, tBodyAccY_mean = V2, tBodyAccZ_mean = V3,
             tBodyAccX_sd = V4, tBodyAccY_sd = V5, tBodyAccZ_sd = V6, 
             tGravityAccX_mean = V41, tGravityAccY_mean = V42, 
             tGravityAccZ_mean = V43, tGravityAccX_sd = V44, 
             tGravityAccY_sd = V45, tGravityAccZ_sd = V46, 
             tBodyAccJerkX_mean = V81, tBodyAccJerkY_mean = V82, 
             tBodyAccJerkZ_mean = V83, tBodyAccJerkX_sd = V84, 
             tBodyAccJerkY_sd = V85, tBodyAccJerkZ_sd = V86, 
             tBodyGyroX_mean = V121, tBodyGyroY_mean = V122, 
             tBodyGyroZ_mean = V123, tBodyGyroX_sd = V124, tBodyGyroY_sd = V125,
             tBodyGyroZ_sd = V126, tBodyGyroJerkX_mean = V161, 
             tBodyGyroJerkY_mean = V162, tBodyGyroJerkZ_mean = V163, 
             tBodyGyroJerkX_sd = V164, tBodyGyroJerkY_sd = V165,
             tBodyGyroJerkZ_sd = V166, tBodyAccMag_mean = V201, 
             tBodyAccMag_sd = V202, tGravityAccMag_mean = V214, 
             tGravityAccMag_sd = V215, tBodyAccJerkMag_mean = V227, 
             tBodyAccJerkMag_sd = V228, tBodyGyroMag_mean = V240, 
             tBodyGyroMag_sd = V241, tBodyGyroJerkMag_mean = V253, 
             tBodyGyroJerkMag_sd = V254, fBodyAccX_mean = V266, 
             fBodyAccY_mean = V267, fBodyAccZ_mean = V268,
             fBodyAccX_sd = V269, fBodyAccY_sd = V270, fBodyAccZ_sd = V271, 
             fBodyAccJerkX_mean = V345, fBodyAccJerkY_mean = V346, 
             fBodyAccJerkZ_mean = V347, fBodyAccJerkX_sd = V348, 
             fBodyAccJerkY_sd = V349, fBodyAccJerkZ_sd = V350, 
             fBodyGyroX_mean = V424, fBodyGyroY_mean = V425, 
             fBodyGyroZ_mean = V426, fBodyGyroX_sd = V427, fBodyGyroY_sd = V428,
             fBodyGyroZ_sd = V429, fBodyAccMag_mean = V503, 
             fBodyAccMag_sd = V504, fBodyAccJerkMag_mean = V516, 
             fBodyAccJerkMag_sd = V517, fBodyGyroMag_mean = V529, 
             fBodyGyroMag_sd = V530, fBodyGyroJerkMag_mean = V542, 
             fBodyGyroJerkMag_sd = V543 )
```
## Step 05: From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
This part follows from Step 04. *colMeans* function is applied to get the mean values of the measurements from the *data* set. And then they are populated in the **newdata.csv** which is saved using *write.table* function.

The new column names for the average of the measurements incluses the prefix *AVG* to all the column names of the measurements.
```{r}
subjectID <- rep(1:30, each= 6)
activityID <- rep (1:6, 30)

newdata <- data.frame(matrix(ncol = 68, nrow=180))
newdata$X1 <- subjectID
newdata$X2 <- activityID

newdata$X2[newdata$X2 == 1] <- "Walking"
newdata$X2[newdata$X2 == 2] <- "WalkingUpStairs"
newdata$X2[newdata$X2 == 3] <- "WalkingDownStairs"
newdata$X2[newdata$X2 == 4] <- "Sitting"
newdata$X2[newdata$X2 == 5] <- "Standing"
newdata$X2[newdata$X2 == 6] <- "Laying"

newdata <- rename(newdata, subjID = X1)
newdata <- rename(newdata, activityID = X2)


## a. Calculating averages and populating the new dataset with its values
for (i in 1:30){
    SubjWalking <- filter(data, subjID == i & activityID == "Walking")
    SubjWalkingAVG <- colMeans(SubjWalking[4:69], na.rm=T)
    newdata[1 + 6*(i-1),3:68] <- SubjWalkingAVG
    
    SubjWalkingUpStairs <- filter(data, subjID == i & activityID == "WalkingUpStairs")
    SubjWalkingUpStairsAVG <- colMeans(SubjWalkingUpStairs[4:69], na.rm=T)
    newdata[2 + 6*(i-1),3:68] <- SubjWalkingUpStairsAVG
    
    SubjWalkingDownStairs <- filter(data, subjID == i & activityID == "WalkingDownStairs")
    SubjWalkingDownStairsAVG <- colMeans(SubjWalkingDownStairs[4:69], na.rm=T)
    newdata[3 + 6*(i-1),3:68] <- SubjWalkingDownStairsAVG
    
    SubjSitting <- filter(data, subjID == i & activityID == "Sitting")
    SubjSittingAVG <- colMeans(SubjSitting[4:69], na.rm=T)
    newdata[4 + 6*(i-1),3:68] <- SubjSittingAVG
    
    SubjStanding <- filter(data, subjID == i & activityID == "Standing")
    SubjStandingAVG <- colMeans(SubjStanding[4:69], na.rm=T)
    newdata[5 + 6*(i-1),3:68] <- SubjStandingAVG
    
    SubjLaying <- filter(data, subjID == i & activityID == "Laying")
    SubjLayingAVG <- colMeans(SubjLaying[4:69], na.rm=T)
    newdata[6 + 6*(i-1),3:68] <- SubjLayingAVG
}


newdata <- rename(newdata, AVGtBodyAccX_mean = X3)
newdata <- rename(newdata, AVGtBodyAccY_mean = X4)
newdata <- rename(newdata, AVGtBodyAccZ_mean = X5)
newdata <- rename(newdata, AVGtBodyAccX_sd = X6)
newdata <- rename(newdata, AVGtBodyAccY_sd = X7)
newdata <- rename(newdata, AVGtBodyAccZ_sd = X8)
newdata <- rename(newdata, AVGtGravityAccX_mean = X9)
newdata <- rename(newdata, AVGtGravityAccY_mean = X10)
newdata <- rename(newdata, AVGtGravityAccZ_mean = X11)
newdata <- rename(newdata, AVGtGravityAccX_sd = X12)
newdata <- rename(newdata, AVGtGravityAccY_sd = X13)
newdata <- rename(newdata, AVGtGravityAccZ_sd = X14)
newdata <- rename(newdata, AVGtBodyAccJerkX_mean = X15)
newdata <- rename(newdata, AVGtBodyAccJerkY_mean = X16)
newdata <- rename(newdata, AVGtBodyAccJerkZ_mean = X17)
newdata <- rename(newdata, AVGtBodyAccJerkX_sd = X18)
newdata <- rename(newdata, AVGtBodyAccJerkY_sd = X19)
newdata <- rename(newdata, AVGtBodyAccJerkZ_sd =X20)
newdata <- rename(newdata, AVGtBodyGyroX_mean = X21)
newdata <- rename(newdata, AVGtBodyGyroY_mean = X22)
newdata <- rename(newdata, AVGtBodyGyroZ_mean = X23)
newdata <- rename(newdata, AVGtBodyGyroX_sd = X24)
newdata <- rename(newdata, AVGtBodyGyroY_sd = X25)
newdata <- rename(newdata, AVGtBodyGyroZ_sd = X26)
newdata <- rename(newdata, AVGtBodyGyroJerkX_mean = X27)
newdata <- rename(newdata, AVGtBodyGyroJerkY_mean  = X28)
newdata <- rename(newdata, AVGtBodyGyroJerkZ_mean = X29)
newdata <- rename(newdata, AVGtBodyGyroJerkX_sd = X30)
newdata <- rename(newdata, AVGtBodyGyroJerkY_sd = X31)
newdata <- rename(newdata, AVGtBodyGyroJerkZ_sd = X32)
newdata <- rename(newdata, AVGtBodyAccMag_mean  = X33)
newdata <- rename(newdata, AVGtBodyAccMag_sd = X34)
newdata <- rename(newdata, AVGtGravityAccMag_mean = X35)
newdata <- rename(newdata, AVGtGravityAccMag_sd = X36)
newdata <- rename(newdata, AVGtBodyAccJerkMag_mean = X37)
newdata <- rename(newdata, AVGtBodyAccJerkMag_sd = X38)
newdata <- rename(newdata, AVGtBodyGyroMag_mean = X39)
newdata <- rename(newdata, AVGtBodyGyroMag_sd = X40)
newdata <- rename(newdata, AVGtBodyGyroJerkMag_mean = X41)
newdata <- rename(newdata, AVGtBodyGyroJerkMag_sd = X42)
newdata <- rename(newdata, AVGfBodyAccX_mean= X43)
newdata <- rename(newdata, AVGfBodyAccY_mean= X44)
newdata <- rename(newdata, AVGfBodyAccZ_mean = X45)
newdata <- rename(newdata, AVGfBodyAccX_sd = X46)
newdata <- rename(newdata, AVGfBodyAccY_sd= X47)
newdata <- rename(newdata, AVGfBodyAccZ_sd = X48)
newdata <- rename(newdata, AVGfBodyAccJerkX_mean = X49)
newdata <- rename(newdata, AVGfBodyAccJerkY_mean = X50)
newdata <- rename(newdata, AVGfBodyAccJerkZ_mean = X51)
newdata <- rename(newdata, AVGfBodyAccJerkX_sd = X52)
newdata <- rename(newdata, AVGfBodyAccJerkY_sd = X53)
newdata <- rename(newdata, AVGfBodyAccJerkZ_sd = X54)
newdata <- rename(newdata, AVGfBodyGyroX_mean = X55)
newdata <- rename(newdata, AVGfBodyGyroY_mean = X56)
newdata <- rename(newdata, AVGfBodyGyroZ_mean = X57)
newdata <- rename(newdata, AVGfBodyGyroX_sd = X58)
newdata <- rename(newdata, AVGfBodyGyroY_sd = X59)
newdata <- rename(newdata, AVGfBodyGyroZ_sd = X60)
newdata <- rename(newdata, AVGfBodyAccMag_mean = X61)
newdata <- rename(newdata, AVGfBodyAccMag_sd  = X62)
newdata <- rename(newdata, AVGfBodyAccJerkMag_mean= X63)
newdata <- rename(newdata, AVGfBodyAccJerkMag_sd= X64)
newdata <- rename(newdata, AVGfBodyGyroMag_mean = X65)
newdata <- rename(newdata, AVGfBodyGyroMag_sd = X66)
newdata <- rename(newdata, AVGfBodyGyroJerkMag_mean =X67)
newdata <- rename(newdata, AVGfBodyGyroJerkMag_sd = X68)
```

The new tidy data is saved as *newdata.csv*
